---
output: github_document
editor_options: 
  chunk_output_type: console
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)

library(saltbush)
```

<!-- badges: start -->
[![](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)
[![R-CMD-check](https://github.com/traitecoevo/saltbush/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/traitecoevo/saltbush/actions/workflows/R-CMD-check.yaml)
[![Codecov test coverage](https://codecov.io/gh/traitecoevo/saltbush/graph/badge.svg)](https://app.codecov.io/gh/traitecoevo/saltbush)
<!-- badges: end -->
# saltbush

*saltbush* processes drone imagery and ausplot vegetation survey data to calculate spectral + taxonomic diversity values for assessment of the 'spectral variability hypothesis'.

# Installation
```{r install, eval= FALSE}
 install.packages("remotes")
 remotes::install_github("traitecoevo/saltbush")
```

Spectral diversity metrics:
+ co-efficient of variance (CV)
+ spectral variance (SV)
+ convex hull volume (CHV)

Taxonomic diversity metrics:
+ species richness
+ shannon's diversity index
+ simpson's diversity index
+ pielou's evenness
+ exponential shannon's index
+ inverse simpson's index

