---
title: "Getting started with saltbush"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting started with saltbush}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "vignette-",
  out.width = "90%"
)
library(saltbush)
```

This vignette walks through a complete spectral-diversity workflow using
only data shipped with the package, so it runs offline. The workflow has
two halves that you eventually combine:

1. **From-the-sky:** turn drone imagery into spectral-diversity metrics
   per area-of-interest (AOI).
2. **From-the-ground:** turn AusPlots point-intercept data into
   taxonomic-diversity metrics per site.

## 1. Build a multiband image from single-band TIFs

The example data ships with five single-band TIFs from a drone flight
over Fowlers Gap, NSW (`blue`, `green`, `red`, `red_edge`, `nir`).

```{r build-multiband, message = FALSE, results = "hide"}
input_dir <- system.file("extdata/create_multiband_image",
                         package = "saltbush")

multiband <- create_multiband_image(
  input_dir,
  desired_band_order = c("blue", "green", "red", "red_edge", "nir"),
  output_dir = tempdir(),
  return_raster = TRUE
)
```

## 2. Extract pixel values inside areas of interest

We then combine the multiband image with an area-of-interest shapefile.
The package ships a small aggregated raster and matching AOI for
demonstration:

```{r extract-pixels}
raster_files <- list.files(
  system.file("extdata/example", package = "saltbush"),
  pattern = "\\.tif$", full.names = TRUE
)

aoi_files <- list.files(
  system.file("extdata/aoi", package = "saltbush"),
  pattern = "NSABHC0009_aoi\\.shp$", full.names = TRUE
)

pixel_values <- extract_pixel_values(
  raster_files, aoi_files,
  c("blue", "green", "red", "red_edge", "nir")
)
head(pixel_values)
```

## 3. Calculate spectral metrics

`calculate_spectral_metrics()` returns three metrics per AOI:

- **CV** — coefficient of variation
- **SV** — spectral variance
- **CHV** — convex hull volume

```{r spectral-metrics}
metrics <- calculate_spectral_metrics(
  pixel_values,
  wavelengths = c("blue", "green", "red", "red_edge", "nir"),
  masked = FALSE,
  rarefaction = FALSE
)
metrics
```

## 4. Compute on-the-ground taxonomic diversity

For the from-the-ground side, the package ships a cached AusPlots
point-intercept dataset for the same site (`NSABHC0009`) so the worked
example does not need network access:

```{r field-diversity}
field_diversity <- calculate_field_diversity(ausplots_NSABHC0009)
field_diversity$taxonomic_diversity
```

The returned metrics are species richness, Shannon's diversity,
Simpson's diversity, Pielou's evenness, exponential Shannon's, and
inverse Simpson's.

To pull a fresh query from the AusPlots database (or to query other
plots), use `ausplotsR::get_ausplots()`:

```{r ausplotsR, eval = FALSE}
my.data <- ausplotsR::get_ausplots(
  my.Plot_IDs = c("NSABHC0009"), veg.PI = TRUE
)
calculate_field_diversity(my.data$veg.PI)
```

## 5. Comparing the two

High-resolution drone images are very large; the example data shipped
with the package is intentionally small (just enough to run examples
and tests). To compare spectral and taxonomic diversity meaningfully,
run the above steps on the full-resolution imagery for your own sites.
